---
title: "StatisticalProject1"
author: "Ashutosh Singh"
date: "26 January 2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
# Statistical Inference Course Project 1

## Overview
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. 

## Simulations

```{r}
# load neccesary libraries
library(ggplot2)
# set constants
lambda <- 0.2 # lambda for rexp
n <- 40 # number of exponetials
num_Simulations <- 1000 # number of Simulations
# set the seed to create reproducability
set.seed(11111)
# run the test resulting in n x numberOfSimulations matrix
exponentialDistributions <- matrix(data=rexp(n * num_Simulations, lambda), nrow=num_Simulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))
```

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))
```

## Sample Mean versus Theoretical Mean

The expected mean $\mu$ of a exponential distribution of rate $\lambda$ is 

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
mu
```

Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
meanOfMeans <- mean(exponentialDistributionMeans$means)
meanOfMeans
```

From the above observations we can see that the expected mean and the avarage sample mean are very close 


## Sample Variance versus Theoretical Variance

The expected standard deviation $\sigma$ of a exponential distribution of rate $\lambda$ is 

$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 

The e

```{r}
stdDev <- 1/lambda/sqrt(n)
stdDev
```

The variance $Var$ of standard deviation $\sigma$ is

$Var = \sigma^2$ 

```{r}
Var <- stdDev^2
Var
```

Let $Var_x$ be the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $\sigma_x$ the corresponding standard deviation.
```{r}
sd_x <- sd(exponentialDistributionMeans$means)
sd_x
Var_x <- var(exponentialDistributionMeans$means)
Var_x
```

As you can see the standard deviations are very close
Since variance is the square of the standard deviations, minor differnces will be more but they are still close to each other.

## Distribution

Comparing the population means & standard deviation with a normal distribution of the expected values.

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = mu , sd = sd), colour = "red", size=1) + 
  geom_vline(xintercept = mu, size=1, colour="#FF0000") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = meanOfMeans, size=1, colour="#0000FF") + 
  scale_x_continuous(breaks=seq(mu-3,mu+3,1), limits=c(mu-3,mu+3)) 
```

From the above graph, we can see that the calculated distribution of means of random sampled exponential distributions overlaps quite nicely with the normal distribution with the expected values based on the given lamba.
